Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy

### 2020-10-06 manuscript.Rmd
adding the dance of the bf

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.44$, meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually $0.36$). In other words, the chances of detecting the effect (was the effect actually non-null) were less than obtaining a head in a coin flip.

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, computing a BF (i.e., comparing two models) does not solve *at all* the problem of low power (we discuss this issue in more depth below). Second, few details are given with regards to the exact models that were compared. Third, and most importantly, the BFs indicate moderate evidence in favour of the null for the OOI and OOS muscles. More precisely, these BFs indicated that the (observed) data are $1 / 0.232 \approx 4.31$ times more likely under the null than under the alternative hypothesis for the OOS and $1 / 0.278 \approx 3.6$ times more likely under the null than under the alternative hypothesis for the OOI. In other words, the evidence is favour of the null is relatively weak and sensitivity analyses (i.e., reporting the BF with varying prior scales for the alternative hypothesis) may unsurprisingly results in various BFs. For instance...

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the *relative evidence* for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this, we present below the results of a simulation aiming to assess the relation between the sample size and the value of the BF.

In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by @moffatt_inner_2020. Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

>>>>

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.44$, meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually $0.36$).

In other words, the chances of detecting the effect (was the effect actually non-null) were less than obtaining a head in a coin flip.

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the *relative evidence* for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we present below the results of two simulations.

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborcayk et al., 2017)."}
# https://julianquandt.com/post/power-analysis-by-data-simulation-in-r-part-ii/
# http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

# sample size
n <- 26

# number of simulated experiments
nsims <- 1e3

# initialises an empty array to store results
results <- data.frame(
  BF10 = numeric(nsims),
  BF10_null = numeric(nsims),
  BF10_double = numeric(nsims)
  )

for (i in 1:nsims) { # for each simulated experiment

  # produce N simulated participants
  x <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y <- rnorm(n = n, mean = 0 + cohensd, sd = 1)

  # produce N simulated participants
  y_null <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y_double <- rnorm(n = n, mean = 0 + cohensd * 2, sd = 1)

  results$BF10[i] <- ttestBF(
    x = x, y = y, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_null[i] <- ttestBF(
    x = x, y = y_null, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_double[i] <- ttestBF(
    x = x, y = y_double, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

results %>%
  pivot_longer(cols = BF10:BF10_double, names_to = "type") %>%
  mutate(
    type = factor(
      type,
      levels = c("BF10_null", "BF10", "BF10_double"),
      labels = c("Cohen's d = 0", "Cohen's d = 0.36", "Cohen's d = 0.72")
      )
    ) %>%
  ggplot(aes(x = log(value), fill = type) ) +
  geom_density(alpha = 0.5, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = 2) +
  theme_bw(base_size = 12) +
  facet_wrap(~ type, scales = "free_x") +
  labs(x = "Natural logarithm of the BF10", y = "Density") +
  scale_colour_brewer(palette = "Dark2", direction = 1) +
  scale_fill_brewer(palette = "Dark2", direction = 1)

prop_d0 <- mean(results$BF10_null > 1)
prop_d0.36 <- mean(results$BF10 > 1)
prop_d0.72 <- mean(results$BF10_double > 1)
```

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis), an effect size of d = 0.36 [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of d = 0.72 [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis (i.e., when the "true" effect size is d = 0), `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis. When the "true" effect size is of d = 0.36, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of d = 0.72, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by @moffatt_inner_2020. Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

### 2020-10-06 manuscript.tex
adding the dance of the bf

Wordcount (excluding abstract, references, tables, and figures): 4306

@@

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.44\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually \(0.36\)). In other words, the chances of detecting the effect (was the effect actually non-null) were less than obtaining a head in a coin flip.

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this, we present below the results of a simulation aiming to assess the relation between the sample size and the value of the BF.

In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis (\(\text{BF}_{10}\)) using the \texttt{BayesFactor} package (Morey \& Rouder, \protect\hyperlink{ref-R-BayesFactor}{2018}). We used a ``medium'' prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with \(1000\) simulations (i.e., 1000 simulated datasets) for each sample size.

>>>>

Wordcount (excluding abstract, references, tables, and figures): 4458

@@

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.44\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually \(0.36\)).

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we present below the results of two simulations.

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/bf-dance-1}

\caption{Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborcayk et al., 2017).}\label{fig:bf-dance}
\end{figure}

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis), an effect size of d = 0.36 (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of d = 0.72 (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}). As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis (i.e., when the ``true'' effect size is d = 0), 6.20\% of the computed BFs are above 0 and hence support the alternative hypothesis. When the ``true'' effect size is of d = 0.36, 71.80\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of d = 0.72, 26.30\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis (\(\text{BF}_{10}\)) using the \texttt{BayesFactor} package (Morey \& Rouder, \protect\hyperlink{ref-R-BayesFactor}{2018}). We used a ``medium'' prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with \(1000\) simulations (i.e., 1000 simulated datasets) for each sample size.
